\pkg{DNest4}: Diffusive Nested Sampling in \proglang{C++} and \proglang{Python}

### 2016-06-08 paper/ms.tex
conditional prior class stuff

We now describe the specific prior distributions we used in this example.
The prior for $N$ was
\begin{align}
p(N) \propto \frac{1}{N+1}
\end{align}
for $N \in \{1, 2, ..., 100\}$.
For the conditional prior of the gaussian parameters,
we used Laplace (biexponential)
distributions\footnote{A Laplace
distribution with location parameter $a$ and scale parameter
$b$ has density $p(x | a, b) = \frac{1}{2}\exp\left(-\frac{1}{b}|x - a|\right)$.}.
Normal distributions would be more convenional, but the analytic cumulative
distribution function (CDF) of the Laplace distribution makes it easier to
implement. These were:
\begin{align}
\mu_j &\sim \textnormal{Laplace}(a_\mu, b_\mu)\\
\ln\sigma_j &\sim \textnormal{Laplace}(a_{\ln\sigma}, b_{\ln\sigma})\\
\ln W_j &\sim \textnormal{Laplace}(0, b_{\ln W}).
\end{align}
The location parameters are denoted with $a$ and scale parameters with $b$.
These priors express the idea that the centers, widths, and relative weights
of the gaussian mixture components are probably clustered around some typical
value.

@@

\subsection{Conditional priors}
The \code{RJObject} class is used to manage the components. An additional
class is needed to define and manage the hyperparameters $\alpha$ (i.e.
define how they are generated from the prior, how they are proposed, and
so on). The and the iid
conditional prior $p(x_i | \alpha)$.
In the example, the class is called \code{MyConditionalPrior} and
is inherited from an abstract base class \code{DNest4::ConditionalPrior}.

The member functions \code{from\_prior}, \code{perturb\_hyperparameters}, and
\code{print} do the same things as the similar functions in a standard model
class, but for the hyperparameters $\alpha$. In addition, three functions
\code{from\_uniform}, \code{to\_uniform}, and \code{log\_pdf}
together define the form of the conditional prior for the
component parameters, $p(x_i | \alpha)$.

>>>>

We now describe the specific prior distributions we used in this example.
The prior for $N$ was
\begin{align}
p(N) \propto \frac{1}{N+1}
\end{align}
for $N \in \{1, 2, ..., 100\}$.
For the conditional prior of the gaussian parameters,
we used Laplace (biexponential)
distributions\footnote{A Laplace
distribution with location parameter $a$ and scale parameter
$b$ has density $p(x | a, b) = \frac{1}{2}\exp\left(-\frac{1}{b}|x - a|\right)$.}.
Normal distributions would be more convenional, but the analytic cumulative
distribution function (CDF) of the Laplace distribution makes it easier to
implement. These were:
\begin{align}
\mu_j &\sim \textnormal{Laplace}(a_\mu, b_\mu)\label{eqn:cond_prior1}\\
\ln\sigma_j &\sim \textnormal{Laplace}(a_{\ln\sigma}, b_{\ln\sigma})\label{eqn:cond_prior2}\\
\ln W_j &\sim \textnormal{Laplace}(0, b_{\ln W}).\label{eqn:cond_prior3}
\end{align}
The location parameters are denoted with $a$ and scale parameters with $b$.
These priors express the idea that the centers, widths, and relative weights
of the gaussian mixture components are probably clustered around some typical
value.

@@

\subsection{Conditional priors}
The \code{RJObject} class is used to manage the components. An additional
class is needed to define and manage the hyperparameters $\alpha$ (i.e.
define how they are generated from the prior, how they are proposed, and
so on). The iid
conditional prior $p(x_i | \alpha)$ is also defined by this extra class.
In the example, the class is called \code{MyConditionalPrior} and
is inherited from an abstract base class \code{DNest4::ConditionalPrior}
which insists that certain member functions be specified.

The member functions \code{from\_prior}, \code{perturb\_hyperparameters}, and
\code{print} do the same things as the similar functions in a standard model
class, but for the hyperparameters $\alpha$. 
In addition, three functions
\code{from\_uniform}, \code{to\_uniform}, and \code{log\_pdf}
together define the form of the conditional prior for the
component parameters, $p(x_i | \alpha)$. Each of these takes a
\code{std::vector} of \code{double}s by reference as input.

The \code{log\_pdf} function just evaluates $\ln p(x_i | \alpha)$
at the current value of $\alpha$. The position $x_i$ where this is evaluated
is passed in via the input vector. In the example, these
were defined using Equations~\ref{eqn:cond_prior1}--\ref{eqn:cond_prior3}.
To simplify this, we have defined a class for Laplace distributions,
rather than explicitly writing out the densities. We decided (arbitarily)
that parameters 0,1, and 2 are $\mu$, $\ln(\sigma)$, and $W$ respectively.
\begin{CodeChunk}
\begin{CodeInput}
// vec = {mu, log_sigma, log_weight}
double MyConditionalPrior::log_pdf(const std::vector<double>& vec) const
{
    // Three Laplace distributions
    Laplace l1(location_mu, scale_mu);
    Laplace l2(location_log_sigma, scale_log_sigma);
    Laplace l3(0.0, scale_log_weight);
	return l1.log_pdf(vec[0]) + l2.log_pdf(vec[1]) + l3.log_pdf(vec[2]);
}
\end{CodeInput}
\end{CodeChunk}

The functions \code{to\_uniform} and \code{from\_uniform} must implement
the cumulative distribution function (CDF) of the conditional
prior and its inverse. i.e. \code{from\_uniform}, if the input vector contains
iid draws from Uniform(0, 1), these should be modified to become draws from
$p(x|\alpha)$, and \code{to\_uniform} should be the inverse of
\code{from\_uniform}.
Both of these functions modify the vector argument in-place.
For the example, we again used an external class to define these functions
for the Laplace distribution:

\begin{CodeChunk}
\begin{CodeInput}
// vec = {mu, log_sigma, log_weight}
void MyConditionalPrior::from_uniform(std::vector<double>& vec) const
{
    // Three Laplace distributions
    Laplace l1(location_mu, scale_mu);
    Laplace l2(location_log_sigma, scale_log_sigma);
    Laplace l3(0.0, scale_log_weight);

    vec[0] = l1.cdf_inverse(vec[0]);
    vec[1] = l2.cdf_inverse(vec[1]);
    vec[2] = l3.cdf_inverse(vec[2]);
}

// vec = {mu, log_sigma, log_weight}
void MyConditionalPrior::to_uniform(std::vector<double>& vec) const
{
    // Three Laplace distributions
    Laplace l1(location_mu, scale_mu);
    Laplace l2(location_log_sigma, scale_log_sigma);
    Laplace l3(0.0, scale_log_weight);

    vec[0] = l1.cdf(vec[0]);
    vec[1] = l2.cdf(vec[1]);
    vec[2] = l3.cdf(vec[2]);
}
\end{CodeInput}
\end{CodeChunk}
